Red, Blue, or Purple? 2024’s Battleground States Tell a Story

### 2024-11-04 paper/paper.qmd
paper final draft

---
title: "Red, Blue, or Purple? 2024’s Battleground States Tell a Story"
subtitle: "Pennsylvania, Ohio, and Nevada Become Game Changers as Trump Gains Ground"
author: 
  - Yingke He
  - Ziheng Zhong
thanks: "Code and data are available at: [https://github.com/iJustinn/Election_Prediction.git](https://github.com/iJustinn/Election_Prediction.git)."
date: today
date-format: long
abstract: "This paper presents an analysis of polling data for the 2024 U.S. Presidential election, focusing on key candidates, notably Donald Trump and Kamala Harris, to provide an in-depth assessment of the race. Employing a Hierarchical Bayesian Model, the study reveals substantial state-level variations, with Pennsylvania, Ohio, Georgia, and Nevada identified as pivotal battlegrounds that show a tendency to lean toward Trump. These findings emphasize the dynamic and shifting landscape of U.S. politics and underscore the critical impact of swing states on election results."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

This research contributes to the field of election forecasting by integrating aggregated polling data with a robust modeling approach, offering insights for political analysts, campaign strategists, and policymakers. By identifying pivotal swing states and addressing possible polling biases, this study provides stakeholders with tools to anticipate shifts in voter preferences and adapt strategies effectively in a dynamic political landscape (@citeSwing).

@@

The dataset for this analysis is derived from FiveThirtyEight's 2024 National Presidential Polls [@cite538], capturing a broad array of polling data related to the 2024 U.S. Presidential election. This dataset includes survey results from various polling organizations across the country, offering insights into voter preferences by candidate, demographic group, and geographic location. Key variables encompass the polling organization, sample size, polling dates, candidate support percentages, and methodological notes. These elements are essential for interpreting public sentiment over the course of the election campaign, providing snapshots of preferences at distinct points in time. Additionally, variables such as “weight”—applied to adjust for factors like sample size and sample representativeness—enable more precise aggregate predictions. The dataset's context emphasizes its role in informing both the public and analysts about voter behavior trends throughout an election year, underscoring the value of reliable and transparent polling data in building public trust and supporting informed decision-making.

@@

In this study, polling data provides insights into public opinion across various states, capturing candidate support leading up to the 2024 U.S. Presidential election. The data is sourced from multiple polling organizations, each employing distinct methodologies that affect data quality and representativeness. Understanding these methodologies is essential for accurately interpreting the polling data. The primary categories of measurement methods include:

@@

2. **Interactive Voice Response (IVR): **
Pollsters like Trafalgar Group and Rasmussen Reports frequently use IVR, or “robo-polling,” where an automated system asks questions over the phone. IVR is more affordable and efficient for reaching large numbers of people. However, because respondents interact with a machine rather than a live person, they may feel less invested in the survey, impacting response rates and depth of engagement. IVR allows pollsters to obtain timely snapshots of voter intentions, though it may miss out on nuanced responses that a live interviewer could capture.

@@

5. **Specialized In-Person and University-Affiliated Research: **
Some university-affiliated pollsters, such as Quinnipiac University and the University of New Hampshire Survey Center, use face-to-face interviews or demographic-specific focus groups to gain deeper insights. These methods can provide highly detailed responses and enable a thorough understanding of specific voter segments, although they are resource-intensive and generally reserved for more specialized research purposes.

The diversity of data collection methods directly influences how public opinion is measured and interpreted across pollsters. For example, live interviews often capture nuanced voter sentiment but may underrepresent younger demographics who primarily use mobile devices or engage online. In contrast, IVR and online panel methods enable cost-effective, large-scale data collection but may lack depth and introduce self-selection biases. Mixed-mode approaches attempt to address these limitations by incorporating both traditional and digital respondents, potentially resulting in more representative data.

@@

According to @tbl-outcome_variables, the outcome variables include critical components for analyzing polling data, such as State, Pollster, Pollscore, Candidate Name, Percentage (pct), Sample Size, Start Date, End Date, and Days to Election. These variables provide the foundation for estimating actual vote shares by capturing regional differences, pollster credibility, candidate-specific support, and timing in relation to Election Day. Together, they ensure a nuanced approach to election forecasting that accounts for variability in poll methodology and voter sentiment shifts over time.

@@

## Predictor variables 
This study includes several key predictor variables that provide insights into regional voting patterns, polling methodologies, and the timing of voter sentiment shifts. Each variable plays a unique role in forecasting vote shares for the 2024 U.S. Presidential election. Below shown in @tbl-predictor_variables are detailed descriptions of each predictor.

@@

@fig-pollscore_distribution shows the distribution of poll scores. The plot shows two primary clusters, with peaks around \(-1\) and near \(0\). This bimodal distribution suggests that pollsters' ratings might tend to fall into two categories, potentially reflecting different methodologies, biases, or levels of reliability among pollsters. For example, scores clustered around \(-1\) could represent pollsters with a more consistent lean or bias in one direction, while scores around \(0\) may indicate neutral or balanced ratings. These clusters highlight the importance of adjusting for pollster reliability when aggregating data for election predictions, as different pollsters contribute varying levels of accuracy and bias. Incorporating such adjustments can enhance the model’s ability to produce a balanced and accurate forecast, providing a more nuanced view of candidate support that takes into account the quality and tendencies of each poll source.

@@

By incorporating these state-specific adjustments, the model better aligns with local voting behaviors, capturing regional nuances and improving the reliability of its predictions.

@@

@tbl-weighted_prediction illustrates the model’s adjustment of weighted predictions by accounting for recency. This adjustment decreases the influence of older polls while giving more recent polls greater weight, enhancing prediction relevance as Election Day approaches. Here, the "Total Weight" column reflects the combined impact of poll score, sample size, and recency. For example, in Arizona, Donald Trump’s weighted percentage is 48.66 with a significant total weight of -915.07, indicating a substantial recent adjustment due to proximity to the election. Kamala Harris's weighted percentage in the same state is 46.66 with a weight of -716.47, reflecting similar adjustments. This recency factor ensures that the model prioritizes fresher data, which better captures evolving voter sentiment, thereby making predictions more accurate as the election nears.

@@

Model predictions for key states reveals a competitive landscape for the 2024 presidential election, with slight advantages for each candidate in various swing states. Donald Trump is projected to lead in most states, including Arizona, Florida, Michigan, Nevada, Ohio, Texas, and Wisconsin. However, Kamala Harris is predicted to secure a narrow edge in Georgia, North Carolina, and Pennsylvania, reflecting her competitiveness in these critical battlegrounds. This distribution indicates that while Trump may maintain a slight overall advantage in the swing states, Harris still shows potential to contest and even lead in pivotal regions, underscoring the unpredictable nature of the election. The relatively close percentages in each state illustrate the highly competitive environment and the significance of voter turnout and campaign efforts in determining the final outcome.

@@

Incorporating state effects into the model allows for more precise predictions that consider both national trends and unique state-level variations. This approach strengthens the model’s reliability by recognizing and adjusting for regional differences, resulting in predictions that more accurately capture the complexities of voter sentiment across the United States. By accounting for these state-specific effects, the model generates insights that are both detailed and reflective of the country’s diverse political landscape.

@@

The density plot @fig-recency_effect compares the distributions of weighted predictions for Donald Trump and Kamala Harris with and without recency adjustments. For both candidates, predictions are generally concentrated in similar ranges, with Donald Trump’s support mostly clustering around 40-60% and Kamala Harris’s around 30-50%. This similarity in distribution suggests that the recency adjustment does not cause a drastic shift in overall predictions for either candidate. However, some differences are noticeable in the tails of the distributions, particularly for Donald Trump, where predictions with recency adjustments exhibit a slightly tighter distribution, potentially indicating a more stable support level as the election date nears.

@@

Interactive Voice Response (IVR), commonly referred to as robo-polling, allows pollsters to conduct large-scale surveys quickly and affordably by using an automated system to ask questions over the phone. While this method is efficient, it has significant limitations. Respondents may feel less engaged when interacting with a machine, leading to lower response rates and potentially less accurate data. IVR surveys also lack the flexibility for follow-up questions or clarifications, which can result in oversimplified responses that miss the nuances of voter sentiment. This limitation is especially relevant when polling on complex political issues where a more interactive approach could yield deeper insights.

@@

Finally, specialized in-person and university-affiliated research methods, such as those used by Quinnipiac University and the University of New Hampshire Survey Center, offer a detailed and in-depth approach to data collection. Face-to-face interviews and demographic-specific focus groups can yield rich, context-specific insights into voter behavior. However, these methods are resource-intensive and time-consuming, which limits their scalability. They are typically only feasible for small-scale or targeted studies rather than large-scale polling efforts. Moreover, the high cost of conducting in-person interviews makes it challenging to implement this approach across multiple regions or demographics, which restricts its representativeness in broader studies.

@@

This survey aims to capture a comprehensive view of voter sentiments across the United States, focusing on key battleground states such as Pennsylvania, Ohio, and Nevada. The survey will provide insights into voters' preferences, motivations, and key issues that influence their voting behavior. By emphasizing pivotal states and stratifying data by demographics, the survey offers a granular understanding of the political landscape as it relates to the 2024 U.S. Presidential election.

@@

These stratification variables are chosen to capture potential differences in voting preferences and sentiments, allowing a nuanced analysis of voter dynamics in each group.

@@

The primary data source for this study is national polling data on the 2024 U.S. Presidential election, provided by FiveThirtyEight (@cite538). This dataset compiles polling results from a range of organizations, each employing distinct methodologies, sample sizes, and timing intervals. By integrating data from multiple sources, FiveThirtyEight provides a broad view of national public opinion trends, capturing fluctuations in candidate support as the election approaches. Additionally, the dataset includes detailed information on each pollster's historical performance and reliability, supporting a more nuanced evaluation of poll accuracy and potential biases.

@@

Despite these simplifications, the hierarchical Bayesian framework (if applicable) offers flexibility by allowing for uncertainty quantification and incorporating pollster reliability into the model’s predictions. By assigning reliability-based weights to pollsters with stronger track records, the model aims to enhance accuracy, giving greater influence to data sources that have historically demonstrated consistency. This approach not only improves prediction robustness but also provides a framework for capturing the variance across pollster methodologies, enabling a more nuanced and adaptable representation of public opinion as it evolves in the lead-up to the election.

>>>>

---
title: "Red, Blue, or Purple? 2024’s Battleground States Tell a Story"
subtitle: "Pennsylvania, Ohio, and Nevada Become Game Changers as Trump Gains Ground"
author: 
  - Yingke He
  - Ziheng Zhong
thanks: "Code and data are available at: [https://github.com/iJustinn/Election_Prediction.git](https://github.com/iJustinn/Election_Prediction.git)."
date: today
date-format: long
abstract: "This paper presents an analysis of polling data for the 2024 U.S. Presidential election, focusing on key candidates, notably Donald Trump and Kamala Harris, to provide an in-depth assessment of the race. Employing a Hierarchical Bayesian Model, the study shows substantial state-level variations, with Pennsylvania, Ohio, Georgia, and Nevada identified as pivotal battlegrounds that show a tendency to lean toward Trump. These findings emphasize the dynamic and shifting landscape of U.S. politics and underscore the key impact of swing states on election results."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

This research contributes to election forecasting by integrating aggregated polling data with a robust modeling framework, providing valuable resources for political analysts, campaign strategists, and policymakers. By identifying influential swing states and adjusting for potential polling biases, the study offers stakeholders tools to anticipate shifts in voter preferences and adapt strategies effectively within a dynamic political landscape (@citeSwing).

@@

The dataset used in this analysis is derived from FiveThirtyEight's 2024 National Presidential Polls [@cite538], which offers a comprehensive compilation of polling data on the 2024 U.S. Presidential election. This dataset includes survey results from a variety of polling organizations across the nation, presenting detailed information on voter preferences by candidate, demographic group, and geographic area. Key variables include polling organization, sample size, polling dates, candidate support percentages, and methodological annotations, all of which are essential for interpreting public sentiment throughout the campaign period and capturing voter preferences at specific intervals. Additionally, variables such as “weight”—used to adjust for sample size and demographic representativeness—enhance the accuracy of aggregated predictions. The dataset’s extensive coverage serves as a resource for the public and analysts alike, illuminating voter behavior trends throughout the election year and emphasizing the value of transparent polling data in fostering public trust and informed decision-making.

@@

This study uses polling data to assess public opinion across multiple states, tracking candidate support trends leading up to the 2024 U.S. Presidential election. The data is gathered from various polling organizations, each employing unique methodologies that influence both data quality and representativeness. A clear understanding of these methodologies is essential for accurate interpretation of the polling results. The primary measurement methods can be grouped into the following categories:

@@

2. **Interactive Voice Response (IVR): **
Pollsters like Trafalgar Group and Rasmussen Reports frequently use IVR, or “robo-polling,” where an automated system asks questions over the phone. IVR is more affordable and efficient for reaching large numbers of people. However, because respondents interact with a machine rather than a live person, they may feel less invested in the survey, impacting response rates and depth of engagement. IVR allows pollsters to obtain timely snapshots of voter intentions, though it may miss out on responses that a live interviewer could capture.

@@

5. **Specialized In-Person and University-Affiliated Research: **
This study examines polling data to evaluate public opinion across different states, tracking trends in candidate support as the 2024 U.S. Presidential election approaches. The data is sourced from various polling organizations, each employing distinct methodologies that affect data quality and representativeness. Understanding these methodologies is crucial for accurately interpreting the results. The primary measurement methods are categorized as follows:

The diversity of data collection methods directly influences how public opinion is measured and interpreted across pollsters. For example, live interviews often capture voter sentiment but may underrepresent younger demographics who primarily use mobile devices or engage online. In contrast, IVR and online panel methods enable cost-effective, large-scale data collection but may lack depth and introduce self-selection biases. Mixed-mode approaches attempt to address these limitations by incorporating both traditional and digital respondents, potentially resulting in more representative data.

@@

According to @tbl-outcome_variables, the outcome variables include critical components for analyzing polling data, such as State, Pollster, Pollscore, Candidate Name, Percentage (pct), Sample Size, Start Date, End Date, and Days to Election. These variables provide the foundation for estimating actual vote shares by capturing regional differences, pollster credibility, candidate-specific support, and timing in relation to Election Day. Together, they ensure an approach to election forecasting that accounts for variability in poll methodology and voter sentiment shifts over time.

@@

## Predictor variables 
This study incorporates several essential predictor variables that illuminate regional voting patterns, polling methodologies, and the timing of shifts in voter sentiment. Each variable contributes uniquely to forecasting vote shares in the 2024 U.S. Presidential election. Detailed descriptions of each predictor are presented in @tbl-predictor_variables.

@@

@fig-pollscore_distribution shows the distribution of poll scores. The plot shows two primary clusters, with peaks around \(-1\) and near \(0\). This bimodal distribution suggests that pollsters' ratings might tend to fall into two categories, potentially reflecting different methodologies, biases, or levels of reliability among pollsters. For example, scores clustered around \(-1\) could represent pollsters with a more consistent lean or bias in one direction, while scores around \(0\) may indicate neutral or balanced ratings. These clusters highlight the importance of adjusting for pollster reliability when aggregating data for election predictions, as different pollsters contribute varying levels of accuracy and bias. Incorporating such adjustments can enhance the model’s ability to produce a balanced and accurate forecast, providing a view of candidate support that takes into account the quality and tendencies of each poll source.

@@

By incorporating these state-specific adjustments, the model better aligns with local voting behaviors, capturing regional improving the reliability of its predictions.

@@

@tbl-weighted_prediction illustrates the model’s adjustment of weighted predictions by accounting for recency. This adjustment decreases the influence of older polls while giving more recent polls greater weight, enhancing prediction relevance as Election Day approaches. Here, the "Total Weight" column reflects the combined impact of poll score, sample size, and recency. For example, in Arizona, Donald Trump’s weighted percentage is 48.66 with a significant total weight of -915.07, indicating a substantial recent adjustment due to proximity to the election. Kamala Harris's weighted percentage in the same state is 46.66 with a weight of -716.47, reflecting similar adjustments. This recency factor ensures that the model prioritizes newer data, which better captures evolving voter sentiment, thereby making predictions more accurate as the election nears.

@@

Model predictions for key states shows a competitive landscape for the 2024 presidential election, with slight advantages for each candidate in various swing states. Donald Trump is projected to lead in most states, including Arizona, Florida, Michigan, Nevada, Ohio, Texas, and Wisconsin. However, Kamala Harris is predicted to secure a narrow edge in Georgia, North Carolina, and Pennsylvania, reflecting her competitiveness in these critical battlegrounds. This distribution indicates that while Trump may maintain a slight overall advantage in the swing states, Harris still shows potential to contest and even lead in pivotal regions, underscoring the unpredictable nature of the election. The relatively close percentages in each state illustrate the highly competitive environment and the significance of voter turnout and campaign efforts in determining the final outcome.

@@

Incorporating state effects into the model allows for more precise predictions by accounting for both national trends and unique state-level variations. This approach improves the model’s reliability by acknowledging and adjusting for regional differences, leading to predictions that more accurately reflect the complexities of voter sentiment across the United States. By integrating these state-specific factors, the model generates findings that are both detailed and representative of the nation’s varied political landscape.

@@

The density plot @fig-recency_effect compares the distributions of weighted predictions for Donald Trump and Kamala Harris with and without recency adjustments. For both candidates, predictions are generally concentrated in similar ranges, with Donald Trump’s support mostly clustering around 40-60% and Kamala Harris’s around 30-50%. This similarity in distribution suggests that the recency adjustment does not cause a big shift in overall predictions for either candidate. However, some differences are noticeable in the tails of the distributions, particularly for Donald Trump, where predictions with recency adjustments exhibit a slightly tighter distribution, potentially indicating a more stable support level as the election date nears.

@@

Interactive Voice Response (IVR), often referred to as robo-polling, allows researchers to conduct extensive surveys efficiently and at lower costs by using automated telephone systems to administer questions. While this approach enhances survey speed and reach, it has notable limitations. Interaction with an automated system can reduce respondent engagement, potentially lowering response rates and compromising data accuracy. Additionally, IVR lacks the adaptability needed for follow-up questions or clarifications, which often leads to simplified responses that do not capture the complexities of voter sentiment. This limitation is particularly pronounced in surveys on political issues, where a more interactive format would likely produce more detailed and insightful responses.

@@

Specialized in-person and university-affiliated research methods, such as those employed by Quinnipiac University and the University of New Hampshire Survey Center, provide a thorough approach to data collection. Face-to-face interviews and demographic-specific focus groups can yield detailed, context-rich information on voter behavior. However, these methods are resource-intensive and time-consuming, which restricts their scalability. As a result, they are generally suitable only for small-scale or targeted studies rather than for extensive polling efforts. Additionally, the high costs associated with conducting in-person interviews make it difficult to apply this approach across diverse regions or demographic groups, thereby limiting its broader representativeness.

@@

This survey seeks to capture a comprehensive view of voter sentiments across the United States, with a particular focus on key battleground states such as Pennsylvania, Ohio, and Nevada. It aims to assess voters' preferences, motivations, and the central issues influencing their behavior. By concentrating on pivotal states and stratifying data by demographic groups, the survey provides a detailed analysis of the political landscape in relation to the 2024 U.S. Presidential election.

@@

These stratification variables are chosen to capture potential differences in voting preferences and sentiments, allowing a meaningful analysis of voter dynamics in each group.

@@

The primary data source for this study is national polling data on the 2024 U.S. Presidential election, provided by FiveThirtyEight (@cite538). This dataset compiles polling results from a range of organizations, each employing distinct methodologies, sample sizes, and timing intervals. By integrating data from multiple sources, FiveThirtyEight provides a broad view of national public opinion trends, capturing fluctuations in candidate support as the election approaches. Additionally, the dataset includes detailed information on each pollster's historical performance and reliability, supporting a more meaningful evaluation of poll accuracy and potential biases.

@@

Despite these simplifications, the hierarchical Bayesian framework (if applicable) offers flexibility by allowing for uncertainty quantification and incorporating pollster reliability into the model’s predictions. By assigning reliability-based weights to pollsters with stronger track records, the model aims to enhance accuracy, giving greater influence to data sources that have historically demonstrated consistency. This approach not only improves prediction robustness but also provides a framework for capturing the variance across pollster methodologies, enabling a more adaptable representation of public opinion as it evolves in the lead-up to the election.
